Reasoning About a Machine with Local Capabilities

### 2018-04-15 presentations/esop18-04-2018/main.tex
Fix typo.
small changes in presentation
Presentation notes

@@

\begin{frame}[fragile]
  \frametitle{How Do We Reason About Programs Informally}
  % Motivation
  \centering

  \begin{tabular}{c}
    \begin{lstlisting}
let x = ref 0 in
  (*$\lambda$*)f.(x := 0;
      f();
      x := 1;
      f();
      assert(x == 1))
    \end{lstlisting}
  \end{tabular}

  \note{Informally, what does this program do?}
  \note{It allocates a reference to 0 and saves a closure to x. The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
  \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
  \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point. The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.} 
\end{frame}

\begin{frame}
  \frametitle{Traditional Stack Pointers}
  \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
    \end{scope}
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);

@@

    % read other stack frames
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);

    % break well-bracketedness
    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
      skip caller frame};
  \end{tikzpicture}
  \note{We need something to enforce security properties on a low-level machine.}
\end{frame}

@@

  \begin{columns}
\begin{column}{0.5\textwidth}
  \begin{itemize}
  \item Low-level machine
    \note{Has instructions as you would expect, load, store, jmp, etc.}
  \item Capabilities replace pointers
    \begin{itemize}
    \item Pointer
    \item Range of authority
    \item Kind of authority
      \begin{itemize}
      \item read/write/execute
      \item enter
      \end{itemize}
      \note{Has instructions for manipulating capabilities.}
    \end{itemize}
    \item Authority checked dynamically
  \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
    \begin{center}
      % Insert capability figure
     \end{center}
\end{column}
\end{columns}
  \note{This particular setup gives a very fine-grained memory control.}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 1}
  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \begin{onlyenv}<+-+(2)>
      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    \note{Evil callee stores stack pointer on the heap.}
    % stores stack pointer in heap
    \begin{onlyenv}<+->
      \begin{scope}
        \fill[gray!20] (14,6) rectangle (18,13);
        \draw (14,6) -- (14,13);
        \draw (18,6) -- (18,13);
      \end{scope}
      \draw (16,14) node {heap memory};
      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \end{onlyenv}
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
    \end{onlyenv}

    \note{Evil callee returns to the caller.}
    % attacker returns
    \begin{onlyenv}<+>
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
    \end{onlyenv}

    \note{caller uses a bit more of the stack and calls the evil callee again.}
    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
    % attacker gets called again + uses old stack pointer.
    \begin{onlyenv}<+>
      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local Capabilities}
  % Short description of local capabilities
  \begin{itemize}
  \item Capabilities tagged with locality (local or global)   \note{We call non-local capabilities global}
  \item New \emph{write-local} permission.
  \item Local capabilities can only be stored by capabilities with \emph{write-local} permission
  \end{itemize}
  Calling convention highlights
  \begin{itemize}
  \item Stack capability is local with permission read, write-local, and execute.
  \item Clear stack before passing stack capability to untrusted code.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities Prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw node[right=0cm of sp1] { stack pointer };

    % stores stack pointer in heap
    \begin{scope}
      \fill[gray!20] (14,6) rectangle (18,13);
      \draw (14,6) -- (14,13);
      \draw (18,6) -- (18,13);
    \end{scope}
    \draw (16,14) node {heap memory};
    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
\end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % first step: capabilities
    \begin{onlyenv}<+>
      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
    \end{onlyenv}

    % attacker stores copy of stack pointer high in the stack
    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<.-.(2)>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}
    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
    \end{onlyenv}

    % attacker calls trusted code again
%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.-.(1)>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}
    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
\end{frame}

\begin{frame}
  \frametitle{Calling Convention (Continued)}
  \begin{itemize}
  \item $\dots$
  \item Clear stack and non-argument registers before invoking untrusted code.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Stack Clearing Prevents Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart
    % attacker stores copy of stack pointer high in the stack
    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.-.(1)>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

>>>>

\usepackage{pgfpages}
\setbeameroption{show notes}
\setbeameroption{show notes on second screen=right}

@@

\begin{frame}[fragile]
  \frametitle{What Does This Program Do?}
  % Motivation
  \centering

  \begin{tabular}{c}
    \begin{lstlisting}
let x = ref 0 in
  (*$\lambda$*)f.(x := 0;
      f();
      x := 1;
      f();
      assert(!x == 1))
    \end{lstlisting}
  \end{tabular}

  \note<1>[item]{Consider program. Assuming a standard ML semantics we can say what it does.}
  \note<1>[item]{Bind x to freshly allocated reference in a closure that...}
  \note<1>[item]{takes callback f, sets x to 0, calls f, sets x to 1, calls f and finally asserts x points to 1.}
  \note<1>[item]{Note the assumption that when we call f, then we return to a specific program point. This is what we call well-bracketedness and we assume we have this in many programming languages.}
  \note<1>[item]{However, in order to execute this code, we need to compile it to assembly. }
  \note<1>[item]{How is well-bracketedness guaranteed? In particular, how is it guaranteed if f is a piece of code we do not trust (maybe handwritten assembly).}
  \note<2>[item]{\emph{We present a calling convention for capability machines that provide well-bracketedness and local state encapsulation as well as a logical relation that allows us to reason about such programs.}}
  \note<2>[item]{Let's first consider how stack pointers traditionally are handled.}

  % \note{Informally, what does this program do?}
  % \note{It allocates a reference to 0 and saves a closure to x. The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
  % \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
  % \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point. The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.} 
\end{frame}

\begin{frame}
  \frametitle{Traditional Stack Pointers}
  % \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
  \note<1-3>[item]{Simply put, a caller calls a function which}
  \note<1-3>[item]{pushes a new stack frame on the stack the callee uses for its execution. }
  \note<1-3>[item]{When the callee is done, then it returns to the caller by popping its stack frams.}
  \note<4->[item]{If \emph{callee (evil) assembly code with no intention to follow the CC}, then there are multiple ways for them to break things:}
  \note<5->[item]{Read or write directly from or to the caller's stack frame, breaking local-state encapsulation}
  \note<6->[item]{Skip the caller's stack frame and return to one further down breaking well-bracketedness.}
  \note<6->[item]{Clearly we need some kind of low-level enforcement mechanism.}

  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
    \end{scope}
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);

@@

        % read other stack frames
    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.-.(1)> (sp0) edge[->,out=235,in=0] (6,8.25);

    % read other stack frames

    \actadv[+]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);

    % break well-bracketedness
    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
      skip caller frame};
  \end{tikzpicture}
  % \note{We need something to enforce security properties on a low-level machine.}
\end{frame}

@@

  \begin{columns}
\begin{column}{0.5\textwidth}
  \begin{itemize}
  \item<1-> Low-level machine
    \note<1->[item]{Capability machines are low-level machines proposed in the systems community.}
    \note<1->[item]{For instance, the CHERI OS operates on one.}
    \note<1->[item]{Has all the instructions we expect, load, store, jmp, etc.}
    % \note{Has instructions as you would expect, load, store, jmp, etc.}
  \item<2-> Capabilities replace pointers
    \begin{itemize}
    \item<3-> Pointer
    \item<4-> Range of authority
    \item<5-> Kind of authority
      \note<5->[item]{Roughly two kinds of capabilities:}
      \begin{itemize}
      \item read, write, and execute
        \note<5->[item]{Memory capabilities, allows you to do all the standard memory operations.}
      \item enter
        \note<5->[item]{Provides encapsulation mechanism which allows seperation of security domains.}
        \note<5->[item]{Can not be used for anything but jump, when jumped to becomes read/execute.}
      \end{itemize}
    \end{itemize}
    \item<6-> Capability manipulation instructions
    \item<7-> Authority checked dynamically
    \end{itemize}
\end{column}

\begin{column}{0.5\textwidth}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \draw (1,9) node {Memory};
    \draw (-3.5,0) node {\phantom{.}};
    \scope
      \clip (-.1,-.1) rectangle (6.1,8.1);
      \fill[fill=white] (0,0) rectangle (6,8);
      \draw (0,0) -- (0,8);
      \draw (6,0) -- (6,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    \draw<3>  (-3,1.5) edge[black,->,in=180] (0,2.75);

    \draw<4-> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
    (-0.25,1) -- (-0.25,4.5) node[draw=black] (sp1) [black,midway,xshift=-12pt] {};
    \draw<4>  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize\phantom{\textsc{rw}}} (sp1);
    \draw<5->  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize \textsc{rw}} (sp1);
  \end{tikzpicture}
\end{column}
\end{columns}
 % \note{This particular setup gives a very fine-grained memory control.}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 1}
%  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
  \note[item]<1->{Let's see how this changes things: Now the \emph{untrusted code cannot immediately read from the caller stack frame}, because the stack capability does not have authority over that part of memory.}
  \note[item]<1->{There is nothing that prevents the \emph{untrusted code from storing the stk ptr on the heap.}}
  \note[item]<3->{Upon return the \emph{callee regains its stack capability} which has authority over the callee stack frame and everything above. }
  \note[item]<3->{The \emph{caller pushes some important things} on the stack and \emph{calls the untrusted code again}. With a smaller stack pointer.}
  \note[item]<4->{The stack pointer the caller gives the untrusted code cannot be used to access the callee stack frame, but because the untrusted code stored the old stack pointer, it now has access to part of the callee's stack frame.}
  \note[item]<4->{Again breaking local state encapsulation.}
  \note[item]<4->{Need a way to make sure stack pointer is not stored for later use.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \begin{onlyenv}<+-+(2)>
      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node[above right=1cm of sp1, xshift=-0.5cm] (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
    \end{onlyenv}

%    \note{Evil callee stores stack pointer on the heap.}
    % stores stack pointer in heap
    \begin{onlyenv}<+->
      \begin{scope}
        \fill[gray!20] (14,6) rectangle (18,13);
        \draw (14,6) -- (14,13);
        \draw (18,6) -- (18,13);
      \end{scope}
      \draw (16,14) node {heap memory};
      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \end{onlyenv}
    \draw<.> (14,8.25) edge[->,red,thick,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (14,8.25) edge[->,red,thick,bend left] node[left] {\footnotesize \textsc{rw}}(spold);
    \end{onlyenv}

%    \note{Evil callee returns to the caller.}
    % attacker returns
    \begin{onlyenv}<+>
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below right] {\footnotesize \textsc{rw}} (sp1);
      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
    \end{onlyenv}

%    \note{caller uses a bit more of the stack and calls the evil callee again.}
%    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
    % attacker gets called again + uses old stack pointer.
    \begin{onlyenv}<+>
      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local Capabilities}
  % Short description of local capabilities
  CHERI inspired
  \begin{itemize}
  \item Capabilities tagged with locality (local or global)   %\note{We call non-local capabilities global}
  \item New \textit{write-local} permission
  \item Local capabilities can only be stored by capabilities with \textit{write-local} permission
  \end{itemize}
  \note[item]<1->{To revoke a capability, we need to find it in memory which means we need access + need to search the entire memory.}
  \note[item]<1->{Restricted where local capabilities can be stored. restricts where we need to look for a capability.}
  \note[item]<1->{We define a calling convention. In order to prevent attack 1, we do the follwoing.}
  \onslide<2->{Calling convention}
  \begin{itemize}[<2->]
  \item Stack capability is local with permission read, write-local, and execute.
    \note[item]<2->{Local stack capability cannot be stored on the heap. We need to be able to store old stack pointers somewhere, traditionally stack. }
  \item No global write-local capabilities.
    \note[item]<2->{Global write-local capabilities would undermine the entire idea as it would allow local capabilities to be stored indirectly.}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities Prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \note{In the attack from before, when the attacker \emph{attempts to store the stack capability} on the heap, then the \emph{machine checks} that we have the correct authority to perform the operation. \emph{Assuming we only have global capabilities} for the heap, it cannot have write-local authority, due to the assumption on the previous slide, so we try to store the stack capability through a capability that does not have write-local authority, so it fails.}

    % recurrent parts
    \stdstackstart

    % attacker...
    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw (11,14) node (a) { stack pointer };
    \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);

    % stores stack pointer in heap
    \begin{scope}
      \fill[gray!20] (14,6) rectangle (18,13);
      \draw (14,6) -- (14,13);
      \draw (18,6) -- (18,13);
    \end{scope}
    \draw (16,14) node {heap memory};
    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw (14,8.25) edge[->,red,thick,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
\end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \note[item]{While this prevents attack 1, we are not quite safe done.}
    % recurrent parts
    \stdstackstart
    \note[item]{Trusted caller calls untrusted code.}
    \note[item]<2->{untrusted code stores the stack pointer on the stack.}
    \note[item]<2->{stack pointer local, but stack pointer has write local permission, so no problem.}
    \note[item]<3->{untrusted code calls some trusted code with a callback.}
    \note[item]<5->{trusted code runs for a bit pushes some local data to the stack and calls the callback.}
    \note[item]<6->{The stack pointer is still on the stack allowing the untrusted code to read write to the stack frame of the trusted code.}
    % first step: capabilities
    \begin{onlyenv}<+>
      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
    \end{onlyenv}

    % attacker stores copy of stack pointer high in the stack
    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<.-.(2)>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);        
    \end{onlyenv}
    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] node[above right,yshift=0.1cm,xshift=-0.2cm] {\footnotesize local,\textsc{rwlx}}(sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (6,13.25) edge[red,thick,->,in=45] node[sloped, above] {\footnotesize local,\textsc{rwlx}}(spold);
    \end{onlyenv}

    % attacker calls trusted code again
%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
        \draw (11,12.5) node (a) { stack pointer };
        \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \end{onlyenv}

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.-.(1)>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};

      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \end{onlyenv}
    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
%  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
\end{frame}

\begin{frame}
  \frametitle{Calling Convention (Continued)}
  \note[item]{Stack is basically the only place we can store local capabiliites.}
  \note[item]{Make sure that untrusted code don't ``sneak'' capabilities between calls on the stack}
  \note[item]{Clear stack and argument registers}
  $\dots$
  \begin{itemize}
  \item Clear stack and non-argument registers before invoking untrusted code.
  \end{itemize}
\end{frame}
